
## Reproducible Research Cource Project 1

 Our goal is to consider the given data set, containing personal activity monitoring device data. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.
 
### Our project consists the next steps:
 1. Loading data
 
```{r}
 our_data_set <- read.csv("activity.csv")
 head(our_data_set)
```
 
 2. Histogram of the total number of steps taken each days
 For this purpose we wil calculate the total number of steps by days
 
 
```{r totalbydays}
 total_steps_by_days <- aggregate(our_data_set$steps ~ our_data_set$date, FUN=sum)
```
 
 
 histogram with mean and median abline building
 
 
```{r totalstepshisto}
 hist(total_steps_by_days[,2], xlab = "Total steps", main = "Total steps by days", prob = TRUE, col = "green")
 abline(v=c(mean(total_steps_by_days[,2]),median(total_steps_by_days[,2])), col=c("blue", "red"), lty=c(1,2), lwd=c(1, 3))
```
 
 
 like we see from our histogram, the median, `r median(total_steps_by_days[,2])`, and mean, `r mean(total_steps_by_days[,2])` are very closed.
 
 4. Time series plot of the average number of steps taken
 
 Average values by time series
 
```{r timeseriesave}
 average_by_intervals <- aggregate(our_data_set$steps ~ our_data_set$interval, FUN=mean)
```
 
 plot building
 
```{r timeseriesplot}
 plot(average_by_intervals[,1], average_by_intervals[,2], type = "l", xlab = "Time Interval", ylab = "Average", main = "Average number of steps by time       series", xlim = c(0, 2355))
```
 
 5. The 5-minute interval that, on average, contains the maximum number of steps
 
```{r maxaverage}
 our_average_max <- max(average_by_intervals[,2])
 our_average_max
```
 
 6. Imputing missing data
 
 First of all we calculate the total number of missing values
 
```{r nasum}
 sum_of_missing <- sum(is.na(our_data_set$steps))
 sum_of_missing
```

 Strategy for imputimg missing data into new dataset. We will try the mean value of time interval without loading additional pakages.
 We use the for loop
 
```{r forloop, cache = TRUE}
 no_na_data_set <- our_data_set
 for(i in 1:nrow(our_data_set)){
    if(is.na(no_na_data_set[i,1])){
        no_na_data_set[i, 1] <- average_by_intervals[average_by_intervals[,1] == no_na_data_set[i,3],2] 
    }
 }
head(no_na_data_set)
```

7. Histogram of the total number of steps taken each day after missing values are imputed

```{r}
 total_steps_no_na_values <- aggregate(no_na_data_set$steps ~ no_na_data_set$date, FUN=sum)
 hist(total_steps_no_na_values[,2], xlab = "Total steps", main = "Total steps by days", prob = TRUE, col = "green")
```
 
 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
 First of all we need to add a new column to dataset, which contains factor variable with two levels - "weekday" and "weekend"
 
```{r}
library(dplyr)
daysOfWeek_dataSet <- mutate(no_na_data_set, dayOfWeek = factor(case_when(
    weekdays(as.Date(date)) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
    weekdays(as.Date(date))%in% c("Sunday", "Saturday")  ~ "weekend")))
head(daysOfWeek_dataSet)
```

```{r}
weekend_subset <- subset(daysOfWeek_dataSet, daysOfWeek_dataSet$dayOfWeek == "weekend")
weekend_average <- aggregate(weekend_subset$steps ~ weekend_subset$interval, FUN = mean)
weekday_subset <- subset(daysOfWeek_dataSet, daysOfWeek_dataSet$dayOfWeek == "weekday")
weekday_average <- aggregate(weekday_subset$steps ~ weekday_subset$interval, FUN = mean)
```

```{r}
par(mfrow = c(1,2))
plot(weekend_average[,1], weekend_average[,2], type = "l", xlab = "Time Interval", ylab = "Weekend Steps Average", main = "Average Steps by time series")
plot(weekday_average[,1], weekday_average[,2], type = "l", xlab = "Time Interval", ylab = "Weekday Steps Average")
```
